Highly Available Transactions: Virtues and Limitations

### 2013-04-29 writeup/hat-appendix.tex
more writing and polishing

\section{Adya Definitions}

@@

\section{Session Definitions}

\begin{definition}[Directly Session-Depends by $S$]
A transaction $T_j$ directly session-depends on transaction $T_i$ if
$T_i$ immediately precedes $T_j$ in the session commit order for session $S$.
\end{definition}

\begin{definition}[Directly Session-Depends]
A transaction $T_j$ directly session-depends on transaction $T_i$ if
$T_j$ directly session-depends on transaction $T_i$ for some session
$S$.
\end{definition}

\section{HAT Guarantees}

\begin{definition}[Read Uncommitted]
\end{definition}

\begin{definition}[Read Committed]
\end{definition}

\begin{definition}[Item Cut Non-Isolation (N-ICI)]
A history $H$ exhibits phenomenon N-ICI if there exists a transaction
$T_i$ such that $T_i$ directly item-depends by $x$ on more than one
other transaction.
\end{definition}

@@

\begin{definition}[Item Cut Isolation (ICI)]
A system that provides Item Cut Isolation prohibits phenomenon N-ICI.
\end{definition}

\begin{definition}[Predicate Cut Non-Isolation (N-ICI)]
A history $H$ exhibits phenomenon N-ICI if there exists a transaction
$T_i$ such that $P= \{p \mid T_i$ directly predicate-read-depends by
$p$ on another other transaction$\}$ and $\bigcap_{p\in P} P \neq
\emptyset$. (TODO: this isn't necessarily it)
\end{definition}

\begin{definition}[Predicate Cut Isolation (ICI)]
A system that provides Predicate Cut Isolation prohibits phenomenon N-ICI.
\end{definition}

\begin{definition}[Non-transactional Atomicity (N-TA)]
A history exhibits phenomenon N-TA if it contains a directed cycle
consisting of exactly one read-dependency edge by $x$ from $T_i$ to
$T_j$ and one anti-dependency edge by $y$ from $T_j$ to $T_i$ and
$T_i$'s corresponding read from $x$ does not precede its corresponding
read from $y$.
\end{definition}

@@

\begin{definition}[Non-monotonic Reads (N-MR)]
A history exhibits phenomenon N-MR if it contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$ with an anti-dependency edge by $i$ from $T_j$ and a
read-dependency edge by $i$ into $T_i$.
\end{definition}

@@

\begin{definition}[Non-monotonic Writes (N-MW)]
A history exhibits phenomenon N-MW if it contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$ and at least one write-dependency edge.
\end{definition}

@@

\begin{definition}[Writes Do Not Follow Reads (N-WFR)]
A history exhibits phenomenon N-WFR if, for all committed transactions
$T_1$, $T_2$, $T_3$ such that $T_2$ write-depends on $T_1$ and $T_3$
write-depends on $T_2$, $T_3$ does not directly anti-depend on $T_1$.
\end{definition}

@@

\begin{definition}[Not-Read Your Writes (N-RYW)]
A history exhibits phenomenon N-RYW if it contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$, at least one anti-dependency edge, and the remainder
anti-dependency or write-dependency edges.
\end{definition}

@@

\begin{observation}
Read Your Writes is not highly available.
\end{observation}

\begin{corollary}
PRAM and Causal Consistency are not highly available.
\end{corollary}

\begin{definition}[Recency Guarantees]
\end{definition}

\begin{definition}[Lost Update]
\end{definition}

\begin{definition}[Write Skew]
\end{definition}

\begin{definition}[Snapshot Isolation]
\end{definition}

\begin{definition}[Repeatable Read]
\end{definition}

\begin{definition}[Safe Register]
\end{definition}

\begin{definition}[Regular Register]
\end{definition}

\begin{definition}[Linearizable Register]
\end{definition}

\begin{definition}[Strong One-copy Serializability]
\end{definition}

\begin{definition}[Model Strength]
We say that a consistency model $M_1$ is stronger than model $M_2$ if
$M_1$ admits a strict subset of the histories admitted by $M_2$.
\end{definition}

\begin{observation}
Transactional Atomicity is weaker than Repeatable Read.
\end{observation}

\begin{observation}
Transactional Atomicity is weaker than Snapshot Isolation.
\end{observation}

\begin{observation}
Causality is weaker than linearizability.
\end{observation}

\begin{observation}
Read Your Writes is weaker than safe register semantics.
\end{observation}

>>>>

\section{Formal Definitions}

In this section, we formally define HAT transactional semantics. Our
formalism is based off of that of Adya~\cite{adya}. For the reader
familiar with his formalism, this is a mostly-straightforward exercise
combining transactional models with distributed systems
semantics. While the novel aspects of this work largely pertain
to \textit{using} these definitions (e.g., in Section~\ref{sec:hats}),
we believe it is instructive to accompany them by appropriate
definitions to both eliminate any ambiguity in prose and for the
enjoyment of more theoretically-inclined readers.

\subsection{Model}

Here, we briefly describe our database model. It is, with the
exception of sessions, identical to that of Adya~\cite{adya}. We omit
a full duplication of his formalism here but highlight several salient
criteria. We refer the interested reader to Adya's Ph.D. thesis,
Section 3.1 (pp. 33--43).

Users submit transactions to a database system that contains multiple
versions of sets of objects. Each transaction is composed of writes,
which create new versions of an object, and reads, which return a
written version or the initial version of the object. A transaction's
last operation is either \texttt{commit} or \texttt{abort}, and there
is exactly one invocation of either these two operations per
transaction. Transactions can either read individual items or read
based on predicates, or logical ranges of data items.

\begin{definition}[Version set of a predicate-based operation]
When a transaction executes a read or write based on a predicate P,
the system selects a version for each tuple in P’s relations. The set
of selected versions is called the \textit{Version set} of this predicate-based
operation and is denoted by Vset(P).
\end{definition}

A history over transactions has two parts: a partial order of events
reflects the ordering of operations with respect to each transaction
and a (total) version order ($\ll$) on the committed versions of each
object.

As a departure from Adya's formalism, to capture the use of session
guarantees, we allow transactions to be grouped into sessions. We
represent sessions as a partial ordering on committed transactions
such that each transaction in the history appears in at most one
session.

\subsection{Conflict and Serialization Graphs}

To reason about isolation anomalies, we use Adya's concept of a
conflict graph, which is composed of dependencies between
transaction. The definitions in this section are directly from Adya,
with two difference. First, we expand Adya's formalism to deal
with \textit{per-item} dependencies. Second, we define session
dependencies (Definitions~\ref{def:sd1} and~\ref{def:sd2}).

@@

\begin{definition}[Directly Session-Depends by $S$]
\label{def:sd1}
A transaction $T_j$ directly session-depends on transaction $T_i$ if
$T_i$ immediately precedes $T_j$ in the session commit order for session $S$.
\end{definition}

\begin{definition}[Directly Session-Depends]
\label{def:sd2}
A transaction $T_j$ directly session-depends on transaction $T_i$ if
$T_j$ directly session-depends on transaction $T_i$ for some session
$S$.
\end{definition}

The dependencies for a history $H$ form a graph called its Directed
Serialization Graph ($DSG(H)$).

\subsection{Transactional Anomalies and Isolation Levels}

Following Adya, we define isolation levels according to
possible \textit{anomalies}--typically represented by cycles in the
serialization graphs. Definitions~\ref{def:n-ici}--\ref{def:lostupdate}
are not found in Adya but are found (albeit not in this formalism) in
Berenson et al.~\cite{ansicritique} and the literature on session
guarantees~\cite{sessionguarantees, vogels-defs}.

\begin{definition}[Write Cycles (G0)]
A history $H$ exhibits phenomenon G0 if $DSG(H)$ contains a directed
cycle consisting entirely of write-dependency edges.
\end{definition}

\begin{definition}[Read Uncommitted]
A system that provides Read Uncommitted isolation prohibits phenomenon G0.
\end{definition}

\begin{definition}[Aborted Reads (G1a)]
A history $H$ exhibits phenomenon G1a if it contains an aborted
transaction $T_1$ and a committed transaction $T_2$ such that $T_2$ has read
some object (maybe via a predicate) modiﬁed by $T_1$.
\end{definition}

\begin{definition}[Intermediate Reads (G1b)]
A history $H$ exhibits phenomenon G1b if it contains a committed
transaction $T_2$ that has read a version of object $x$ (maybe via a
predicate) written by transaction $T_1$ that was not $T_1$’s ﬁnal
modiﬁcation of $x$.
\end{definition}

\begin{definition}[Circular Information Flow (G1c)]
A history $H$ exhibits phenomenon G1c if $DSG(H)$ contains a directed
cycle consisting entirely of dependency edges.
\end{definition}

\begin{definition}[Read Committed]
A system that provides Read Committed isolation prohibits phenomenon G0, G1a, G1b, and G1c.
\end{definition}

\begin{definition}[Item Cut Non-Isolation (N-I-CI)]
\label{def:n-ici}
A history $H$ exhibits phenomenon N-ICI if $DSG(H)$ contains a
transaction $T_i$ such that $T_i$ directly item-depends by $x$ on more
than one other transaction.
\end{definition}

@@

\begin{definition}[Item Cut Isolation (I-CI)]
A system that provides Item Cut Isolation prohibits phenomenon N-I-CI.
\end{definition}

\begin{definition}[Predicate Cut Non-Isolation (N-P-CI)]
A history $H$ exhibits phenomenon N-I-CI if $DSG(H)$ contains a
transaction $T_i$ such that, for two predicate-based reads $p_k$,
$p_j$ in $P$ such that $p_k \subseteq p_j$ and $T_i$ directly
predicate-read-depends by $p_k$ on another transaction $T_l$, but
$T_i$ does not directly predicate-read-depend by $p_j$ on $T_l$.
\end{definition}

\begin{definition}[Predicate Cut Isolation (P-CI)]
A system that provides Predicate Cut Isolation prohibits phenomenon N-P-CI.
\end{definition}

\begin{definition}[Non-transactional Atomicity (N-TA)]
A history $H$ exhibits phenomenon N-TA if $DSG(H)$ contains a directed cycle
consisting of exactly one read-dependency edge by $x$ from $T_i$ to
$T_j$ and one anti-dependency edge by $y$ from $T_j$ to $T_i$ and
$T_i$'s corresponding read from $x$ does not precede its corresponding
read from $y$.
\end{definition}

@@

\begin{definition}[Non-monotonic Reads (N-MR)]
A history $H$ exhibits phenomenon N-MR if $DSG(H)$ contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$ with an anti-dependency edge by $i$ from $T_j$ and a
read-dependency edge by $i$ into $T_i$.
\end{definition}

@@

\begin{definition}[Non-monotonic Writes (N-MW)]
A history $H$ exhibits phenomenon N-MW if $DSG(H)$ contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$ and at least one write-dependency edge.
\end{definition}

@@

\begin{definition}[Writes Do Not Follow Reads (N-WFR)]
A history $H$ exhibits phenomenon N-WFR if, in $DSG(H)$, for all
committed transactions $T_1$, $T_2$, $T_3$ such that $T_2$
write-depends on $T_1$ and $T_3$ write-depends on $T_2$, $T_3$ does
not directly anti-depend on $T_1$.
\end{definition}

@@

\begin{definition}[Not-Read Your Writes (N-RYW)]
A history $H$ exhibits phenomenon N-RYW if $DSG(H)$ contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$, at least one anti-dependency edge, and the remainder
anti-dependency or write-dependency edges.
\end{definition}

@@

\begin{definition}[Lost Update]
A history $H$ exhibits phenomenon Lost if $DSG(H)$ contains a directed
cycle having one or more item-antidependency edges and all edges are
by the same data item $x$.
\label{def:lostupdate}
\end{definition}

\begin{definition}[Write Skew (Adya G2-item)]
A history $H$ exhibits phenomenon Write Skew if $DSG(H)$ contains a directed
cycle having one or more item-antidependency edges.
\end{definition}

For Snapshot Isolation, we depart from Adya's recency-based definition
(see Adya Section 4.3). Nonetheless, implementations of this
definition will still be unavailable due to reliance of preventing
Lost Update.

\begin{definition}[Snapshot Isolation]
A system that provides Snapshot Isolation prevents phenomena G0, G1a,
G1b, G1c, N-P-CI, N-TA, and Lost Update.
\end{definition}

For Repeatable Read, we return to Adya.

\begin{definition}[Repeatable Read]
A system that provides Repeatable Read Isolation prohibits phenomena
G0, G1a, G1b, G1c, and Write Skew.
\end{definition}

For definitions of safe, regular, and linearizable register semantics,
we refer the reader to Herlihy's textbook~\cite{herlihy-art}.

### 2013-04-29 writeup/intro.tex
more writing and polishing

Indeed, serializable transactions---the gold standard of traditional
ACID databases---are not achievable with high availability in the
presence of partitions~\cite{davidson-survey}. However, database
systems have a long tradition of providing weaker isolation and
consistency guarantees~\cite{adya, ansicritique, gray-virtues,
  gray-isolation, kemme-thesis}. Today's ACID and NewSQL databases
often employ weak isolation models due to their concurrency and
performance benefits; weak isolation is overwhelmingly the default
setting in these stores and is often the only option offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics provided by current highly available systems. This raises a
natural question: which of these semantics can be provided while
remaining highly available?

Unfortunately, the relationship between high availability and ACID
semantics has not been well-explored. We have a strong understanding
of weak isolation in the single-server context from which it
originated~\cite{adya, ansicritique, gray-isolation} and many papers
offer techniques for providing distributed
serializability~\cite{bernstein-book, spanner, daudjee-session,
  hstore, krikellas-bargain, calvin, kemme-classification} or snapshot
isolation~\cite{daudjee-snapshot,generalizedsnapshot, kemme-snapshot,
  walter}. Additionally, the distributed computing, programming
languages, and parallel hardware literature contains many
consistency models for single operations on replicated objects~\cite{
  ceri-mechanism, chen-mechanism, pnuts, herlihy-art, eiger, cac,
  sessionguarantees}. However, the literature has largely ignored
semantic guarantees for multiple operations operating on multiple data
items in a highly available distributed environment.

Our main contributions in this paper are as follows. We fill the gap
between the plethora of previously proposed isolation and consistency
models and the goal of high availability.  We classify which among the
wide array of models are achievable with high availability, denoting
them as {\em Highly Available Transactions} (HATs). In doing so, we
demonstrate that although many implementations of HAT semantics are
not highly available, this is an artifact of the implementations
rather than an inherent property of the semantics. Our investigation
shows that, besides serializability, only Snapshot Isolation and
Repeatable Reads are not HAT-compliant, while most other isolation
levels are achievable with high availability. We also demonstrate that
many weak data consistency models from distributed systems are also
both HAT-compliant and simultaneously achievable with traditional ACID
properties.

@@

Our results demonstrate that a wide range of semantic guarantees can
be delivered with availability in face of partitions, but which
guarantees are worthwhile in practice? We study the virtues and
limitations of both classic and HAT systems by surveying practitioner
accounts and research literature, performing experimental analysis on
modern cloud infrastructure, and analyzing representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm, blooml,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints on multi-item updates and sometimes provide limited
uniqueness guarantees. However, particularly for programs with
non-monotonic logic, HATs can fall short, requiring
sometimes-unavailable mechanisms that are more
coordination-intensive. Overall, this paper quantifies the benefits of
Highly Available Transactions can offer as well as the necessary
semantic restrictions they place on applications.

>>>>

Indeed, serializable transactions---the gold standard of traditional
ACID databases---are not achievable with high availability in the
presence of partitions~\cite{davidson-survey}. However, database
systems have a long tradition of providing weaker isolation and
consistency guarantees~\cite{adya, ansicritique, gray-virtues,
  gray-isolation, kemme-thesis}. Today's ACID and NewSQL databases
often employ weak isolation models due to their concurrency and
performance benefits; weak isolation is overwhelmingly the default
setting in these stores and is often the only option offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics provided by current highly available systems. This raises a
natural question: which of these semantics can be provided with high
availability?

Unfortunately, the relationship between ACID semantics and high
availability has not been well-explored. We have a strong
understanding of weak isolation in the single-server context from
which it originated~\cite{adya, ansicritique, gray-isolation} and many
papers offer techniques for providing distributed
serializability~\cite{bernstein-book, spanner, daudjee-session,
  hstore, krikellas-bargain, calvin, kemme-classification} or snapshot
isolation~\cite{daudjee-snapshot,generalizedsnapshot, kemme-snapshot,
  walter}. Additionally, the distributed computing and parallel
hardware literature contains many consistency models for single
operations on replicated objects~\cite{pnuts, herlihy-art, eiger, cac,
  sessionguarantees}. However, the literature lends few clues for
providing semantic guarantees for multiple operations operating on
multiple data items in a highly available distributed environment.

Our main contributions in this paper are as follows. We fill the gap
between the plethora of previously proposed isolation and consistency
models and the goal of high availability.  We classify which among the
wide array of models are achievable with high availability, denoting
them as {\em Highly Available Transactions} (HATs). In doing so, we
demonstrate that although many implementations of HAT semantics are
not highly available, this is an artifact of the implementations
rather than an inherent property of the semantics. Our investigation
shows that, besides serializability, Snapshot Isolation and Repeatable
Reads are not HAT-compliant, while most other isolation levels are
achievable with high availability. We also demonstrate that many weak
data consistency models from distributed systems are also both
HAT-compliant and simultaneously achievable with traditional ACID
properties.

@@

Our results demonstrate that a wide range of semantic guarantees can
be delivered with availability in face of partitions, but which
guarantees are worthwhile in practice? We study the virtues and
limitations of both classic and HAT systems by surveying practitioner
accounts and research literature, performing experimental analysis on
modern cloud infrastructure, and analyzing representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm, crdt}. HAT systems can also enforce arbitrary foreign key
constraints on multi-item updates and sometimes provide limited
uniqueness guarantees. However, particularly for programs with
non-monotonic logic, HATs can fall short, requiring
sometimes-unavailable mechanisms that are more
coordination-intensive. Overall, this paper quantifies the benefits of
Highly Available Transactions can offer as well as the necessary
semantic restrictions they place on applications.

### 2013-04-29 writeup/motivation.tex
some graphs and joe comments
more writing and polishing

\begin{table}
\vspace{.5em}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{graphs/ping-plot.pdf}\vspace{-1em}
\caption{CDF of round-trip times for slowest inter- and intra-
  availability zone links compared to cross-region links.}\vspace{-1em}
\label{fig:rtt}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{small}
\begin{tabular}{|l|c|c|}
\hline
Database & Default & Maximum\\\hline
Actian Ingres 10.0/10S & S & S\\
Aerospike & RC & RC\\
Akiban Persistit & SI & SI\\
Clustrix CLX 4100 & RR & RR\\
Greenplum 4.1 & RC & S \\
IBM DB2 10 for z/OS & CS & S\\
IBM Informix 11.50 & Depends & S\\
MySQL 5.6 & RR & S \\
MemSQL 1b & RC & RC\\
MS SQL Server 2012 & RC & S \\
NuoDB & CR & CR\\
Oracle 11g & RC & SI\\
Oracle Berkeley DB & S & S\\
Oracle Berkeley DB JE & RR & S\\
Postgres 9.2.2 & RC & S\\
SAP HANA & RC & SI\\
ScaleDB 1.02 & RC & RC\\
VoltDB & S & S\\
\hline
\multicolumn{3}{|p{7cm}|}{{\footnotesize RC: read committed, RR: repeatable read, SI: snapshot isolation, S: serializability, CS: cursor stability, CR: consistent read}}\\\hline

\end{tabular}
\caption{Default and maximum isolation levels for ACID and NewSQL
  databases as of January 2013 (from
  \protect\cite{hat-hotos}).}\vspace{-1.5em}
\label{table:existing}
\end{small}
\end{center}
\end{table}

>>>>

\begin{table}[t!]
\vspace{.5em}

@@

\begin{figure}[t!]
\includegraphics[width=\columnwidth]{figs/ping-plot.pdf}\vspace{-1em}
\caption{CDF of round-trip times for slowest inter- and intra-
  availability zone links compared to cross-region links.}\vspace{-1em}
\label{fig:rtt}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{small}
\begin{tabular}{|l|c|c|}
\hline
Database & Default & Maximum\\\hline
Actian Ingres 10.0/10S & S & S\\
Aerospike & RC & RC\\
Akiban Persistit & SI & SI\\
Clustrix CLX 4100 & RR & RR\\
Greenplum 4.1 & RC & S \\
IBM DB2 10 for z/OS & CS & S\\
IBM Informix 11.50 & Depends & S\\
MySQL 5.6 & RR & S \\
MemSQL 1b & RC & RC\\
MS SQL Server 2012 & RC & S \\
NuoDB & CR & CR\\
Oracle 11g & RC & SI\\
Oracle Berkeley DB & S & S\\
Oracle Berkeley DB JE & RR & S\\
Postgres 9.2.2 & RC & S\\
SAP HANA & RC & SI\\
ScaleDB 1.02 & RC & RC\\
VoltDB & S & S\\
\hline
\multicolumn{3}{|p{7cm}|}{{\begin{small}{RC: read committed, RR: repeatable read, SI: snapshot isolation, S: serializability, CS: cursor stability, CR: consistent read}\end{small}}}\\\hline

\end{tabular}
\caption{Default and maximum isolation levels for ACID and NewSQL
  databases as of January 2013 (from
  \protect\cite{hat-hotos}).}\vspace{-2.5em}
\label{table:existing}
\end{small}
\end{center}
\end{table}

### 2013-04-29 writeup/relatedwork.tex
more writing and polishing

\section{Related Work}

@@

Weak consistency and high availability have been well
studied. Serializability has long been known to be
unachievable~\cite{davidson-survey} and Brewer's CAP Theorem has
attracted considerable attention~\cite{gilbert-cap}. Recent work on
PACELC expands CAP by considering connections between ``weak
consistency'' and low latency~\cite{abadi-pacelc}, while several
studies examine weak isolation guarantees~\cite{adya,
  ansicritique}. However, as we have discussed, the connection between
weak, multi-object consistency models in a distributed setting and
high availability is relatively unexplored. With the exception of our
preliminary workshop paper discussing transactional availability and
HAT RC and I-CI~\cite{hat-hotos}--- which this work expands, providing
additional algorithms and impossibility results, a full
taxonomization, and experimental and application analysis---we believe
this paper is the first to explore this area. However, several studies
rigorously classify \textit{non-HAT} semantics: Wisemann et al. have
proposed a three parameter classification for one-copy serializable
database systems, performing a more extensive classification than our
overview in Section~\ref{sec:evaluation}, including eager database
replication techniques~\cite{kemme-classification}. Additionally
several earlier surveys consider distributed mechanisms for achieving
one-copy serializability~\cite{wisemann-survey}, linearizability,
consistent snapshots, and recency bounds~\cite{ceri-mechanism,
  chen-mechanism}.

>>>>

\section{Related Work}
\label{sec:relatedwork}

@@

Weak consistency and high availability have been well
studied. Serializability has long been known to be
unachievable~\cite{davidson-survey} and Brewer's CAP Theorem has
attracted considerable attention~\cite{gilbert-cap}. Recent work on
PACELC expands CAP by considering connections between ``weak
consistency'' and low latency~\cite{abadi-pacelc}, while several
studies examine weak isolation guarantees~\cite{adya,
  ansicritique}. However, as we have discussed, the connection between
weak, multi-object consistency models in a distributed setting and
high availability is relatively unexplored. With the exception of our
preliminary workshop paper discussing transactional availability and
HAT RC and I-CI~\cite{hat-hotos}--- which this work expands, providing
additional algorithms and impossibility results, a full
taxonomization, and experimental and application analysis---we believe
this paper is the first to explore this area. However, several studies
rigorously classify \textit{non-HAT} semantics: Wisemann et al. have
proposed a three parameter classification for one-copy serializable
database systems, performing a more extensive classification than our
overview in Section~\ref{sec:evaluation}, including eager database
replication techniques~\cite{kemme-classification} and distributed
mechanisms for achieving one-copy
serializability~\cite{wisemann-survey}.

%, linearizability, consistent snapshots, and recency
%bounds~\cite{ceri-mechanism, chen-mechanism}.
